---
title: "P8451_HW10_ML"
author: "Ruixi Li"
date: "2024-03-26"
output: html_document
---

```{r library, include=FALSE}
library(tidyverse)
library(caret)
library(rpart.plot)
library(vcd)
library(reshape2)

```

# Research Question

My research question is to generate a hypothesis of which factors among a wide range of prenatal lifestyles and chemical exposures affect birth outcome(measured by birthweight) using data-driven  methods. (Although I propose my research question after conducting EDA, I would put my research question on the top)

# Load .Rdata file and Data Preparation

```{r data preparation, echo=FALSE}
# Load data using path of where file is stored
load("exposome.RData")

# I am planning on exploring the effect of prenatal exposome towards babies' birth weight. So I just removed all postnatal exposure.

codebook|> group_by(period) |> count()
pre = codebook |> filter(period == "Pregnancy") 
name = rownames(pre)
pre_exp = exposome |> select(ID,any_of(name))

#Strip off ID Variable
studydata$ID<-NULL

# summary statistics for all variables
skimr::skim(studydata)# no missing 

# no duplicate
studydata <- unique(studydata) # num of obs not changed, no duplicate

# merge three datasets together
studydata = merge(pre_exp,phenotype,by="ID") |> merge(covariates, by="ID")

```

* There's no missing and duplicate in this dataset. All variables were correctly classified as numeric or factor.In the current analysis we included 87 environmental exposure variables 13 covariates for the pregnancy period and 6 phenotypes were kept although I will only use birthweight(e3_bw) as my outcome. 

# Data Exploration

## Covariates



## Exposome(pre exp)
## categorical variables

```{r cate, warning=FALSE, message=FALSE}
cate = studydata |> select(where(is.factor))
frequency_tables <- cate %>% map(~ as.data.frame(table(.x), .id = "Variable"))
frequency_tables |>knitr::kable()

## Correlations
### Function to calculate Cramér's V
cramers_v <- function(x, y) {
  tbl <- table(x, y)
  chi2 <- chisq.test(tbl, correct = FALSE)$statistic
  n <- sum(tbl)
  phi2 <- chi2 / n
  r <- nrow(tbl) - 1
  k <- ncol(tbl) - 1
  min_rk <- min(r, k)
  v <- sqrt(phi2 / min_rk)
  return(v)
}

### Create an empty matrix to store the results
vars <- names(cate)
n <- length(vars)
cramer_matrix <- matrix(NA, n, n, dimnames = list(vars, vars))

### Calculate Cramér's V for every pair of variables
for (i in 1:n) {
  for (j in 1:n) {
    cramer_matrix[i, j] <- cramers_v(cate[[i]], cate[[j]])
  }
}

### Replace diagonals with 0 for better visualization
diag(cramer_matrix) <- 0

melted_cramer_matrix <- melt(cramer_matrix)

ggplot(melted_cramer_matrix, aes(Var1, Var2, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 0.5, limit = c(0,1), space = "Lab", name="Cramér's V") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
        axis.title = element_blank()) +
  labs(fill = "Cramér's\nV")
# Remove the h_cohort
#studydata = studydata |> select(-h_cohort)


```

## continuous variables

```{r cont}
cont = studydata|> select(where(is.numeric))

summary_table <- cont %>%
  summarise(across(where(is.numeric), list(
    Mean = ~mean(.x, na.rm = TRUE),
    SD = ~sd(.x, na.rm = TRUE),
    Median = ~median(.x, na.rm = TRUE),
    IQR = ~IQR(.x, na.rm = TRUE)
  ))) |>
  pivot_longer(
    cols = everything(), 
    names_to = c(".value", "Statistic"), 
    names_pattern = "(.*)_(.*)"
  ) 
summary_table|> knitr::kable()

# I specify the continuous variables' names in 'name' vector to add h_cohort in this dataset(I want to assess if there's difference between cohorts-especially the exposome)
name_cont = colnames(cont)[-1]
results = studydata |> select(h_cohort, any_of(name))

results_summary = results %>%
  group_by(h_cohort) %>%
  summarize(across(where(is.numeric), 
                   list(mean = ~mean(.x, na.rm = TRUE), 
                        std = ~sd(.x, na.rm = TRUE)))) 

# Conduct anova test for each variable in 'cont' across cohorts


  
## Correlation
## Calculate correlation matrix
cor_matrix <- cor(cont, use = "complete.obs")

### Replace diagonals with 0 for better visualization
diag(cor_matrix) <- 0



# Melt the correlation matrix
melted_cor_matrix <- melt(cor_matrix)

# Plotting
ggplot(melted_cor_matrix, aes(Var1, Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1,1), space = "Lab", 
                       name="Pearson\nCorrelation") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1),
        axis.text.y = element_text(size = 12)) +
  coord_fixed() +
  labs(x = '', y = '', title = 'Correlation Matrix Heatmap')

```

EDA: Since this dataset is a combine data from 6 cohorts,
1. Population characteristics by cohorts
2. Correlation between exposures(overall and within cohort correlation)
3. PCA to reduce dimensionality(meet one of the requirement of constructing a new variable using an unsupervised analysis in option 3-I will ask professor if PCA is considered as creating a new variable; if not, we can also use clustering in the analysis procedure)

pre-processing of exposome(standardizing)




```{r}
# merge three datasets together
studydata = merge(pre_exp,phenotype,by="ID") |> merge(covariates, by="ID")
```


# Implement Elastic Net Algorithm 

```{r partition}
#Partition data for use in demonstration
set.seed(123)
train.indices<-createDataPartition(y=studydata$e3_bw,p=0.7,list=FALSE)
train.data<-studydata[train.indices, ]
test.data<-studydata[-train.indices, ]
```


# Implement Random Forest(or other...)

# ExWAS Analysis

```{r ExWAS}


results <- list()
for (exposure in name) {
    formula <- as.formula(paste("e3_bw ~", exposure))
  
  # Fit the model using the dynamically constructed formula
  model <- glm(formula, data = studydata, family = "gaussian")  
  
  # Store the model's summary in the results list
  results[[exposure]] <- summary(model)
}


```

1. Adjusted association between the exposome and birthweight
2. volcano plots for visualization


